---
title: "STAT40830 Homework 1"
author: "David Whyms"
date: today
date-format: medium
format: pdf
editor: visual
---

## Import data

The data I chose to usefor this submission was the `mtcars` dataset.

```{r}
#| echo: false
#| warning: false
#| label: import_processing

data(mtcars) # import

# two binary variables - converting to factor (helps with plotting later)
mtcars$am<-factor(mtcars$am,levels=c(0,1),labels=c("automatic","manual"))
mtcars$vs<-factor(mtcars$vs,levels=c(0,1),labels=c("v-shaped","straight"))

```

The mtcars dataset contains data on 32 different cars. The dataset features the following 11 variables:

-   **mpg**: Miles/(US) gallon
-   **cyl**: Number of cylinders
-   **disp**: Displacement *(cu.in.)*
-   **hp**: Gross horsepower
-   **drat**: Rear axle ratio
-   **wt**: Weight (1000 lbs)
-   **qsec**: 1/4 mile time
-   **vs**: Engine *(0 = V-shaped, 1 = straight)*
-   **am**: Transmission *(0 = automatic, 1 = manual)*
-   **gear**: Number of forward gears
-   **carb**: Number of carburetors

The following are summary statistics listed for the `mtcars` dataset:

```{r}
#| echo: false
#| warning: false
#| label: summary

summary(mtcars)
```

## Plot

To demonstrate the data, I chose to plot the weight of the car relative to the miles per gallon. The categories are separated by the type of transmission.

```{r}
#| echo: false
#| warning: false
#| label: plot

library(ggplot2) # using ggplot for nicer plot

ggplot(mtcars,aes(x=wt,y=mpg,col=am)) +                       # data
  geom_point(pch=19,size=2) +                                 # scatterplot
  ggtitle("Plot of car miles per gallon vs. weight") +        # add title
  labs(x="weight",y="miles per gallon",col="transmission") +  # axis + legend
  theme(plot.title=element_text(hjust=0.5))      +
  scale_color_brewer(palette="Dark2")# centre title
```

The above plot indicates the following:

-   The relationship between weight and miles per gallon generally decreases. As the weight of a car increases, the miles per gallon is likely to decrease, and vice-versa.

-   Generally, an automatic car tends to have a higher weight than a manual car. As the previous point would imply, this also indicates that an automatic car is likely to have a lower miles-per-gallon than a manual car.
